The Occurrence of Earth-Like Planets Around Other Stars

### 2014-11-26 paper/kepler-selection.tex
Updates to Science format, following Nature rejection :(.

\documentclass{nature}

@@

\bibliographystyle{naturemag}

@@

\input{quantities}

\begin{document}
\title{The Occurrence of Earth-Like Planets Around Other Stars}

\author{Will M. Farr$^{1}$, Ilya Mandel$^{1}$, Chris Aldridge$^{1}$ \& Kirsty Stroud$^{1}$}

@@

\begin{affiliations}
\item School of Physics and Astronomy, University of Birmingham, Birmingham, B15 2TT, United Kingdom
\end{affiliations}

\begin{abstract}
  The quantity $\etaearth$, the number density of planets per star per
  logarithmic planetary radius per logarithmic orbital period at one
  Earth radius and one year period, describes the occurrence of
  Earth-like extrasolar planets.  Measurement of $\etaearth$ is
  complicated by the difficulty of detecting Earth-like planets in
  Earth-like orbits about Sun-like stars.  Here we present
  a measurement of $\etaearth$ from a parameterised forward model of
  the (correlated) period-radius distribution and the observational
  selection function in the most recent (Q17) data release from the
  Kepler satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  Our
  data set comprises 181,568 systems observed under the Kepler
  exoplanet observing program (mostly G-type stars on the main
  sequence\cite{Batalha2010}), producing 2598 planetary candidates.
  We parameterise the distribution of planetary periods and radii
  using a single, correlated Gaussian component; treat selection
  effects using a parameterised transit detection probability based on
  the measured noise level and stellar properties in the Kepler
  catalog; and include an empirically-parameterised, independent
  component in the period-radius distribution to represent
  false-positive planet detections.  Using our model we can
  simultaneously estimate $\etaearth$, place constraints on the planet
  period-radius distribution function, and determine the degree of
  contamination by false-positive candidate identifications.  We find
  $\etaearth = \earange$ (90\% CL).  We conclude that each star hosts
  $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim
  0.2 \REarth$, that the peak of the planet radius distribution lies
  at $\Rpeak = \rpeakrange \REarth$, and that $\ln P$ and $\ln R$ are
  correlated with correlation coefficient $r = \corrcoeffrange$ (all
  90\% CL).  Our empirical model for false-positive contamination is
  consistent with the dominant source being background eclipsing
  binary stars\cite{Fressin2013}, with $\fposrange$ (90\% CL) of the
  candidates being false-positives.  The distribution of planets we
  infer is consistent with a highly-stochastic planet formation
  process producing many correlated, fractional changes in planet
  sizes and orbits.  Our approach of determining both the intrinsic
  distribution of objects and selection effects empirically from
  survey data is generally applicable.
\end{abstract}

@@

The probability that a planet's orbital plane will align with the line-of-sight to
Earth and thereby produce a transit signal is
\begin{equation}
  \label{eq:ptransit}
  p_\mathrm{transit} = 0.0016\, \frac{\RStar}{\RSun}
  \left(\frac{\MStar}{\MSun}\right)^{-1/3} \left(\frac{P}{1\,\mathrm{yr}}\right)^{-2/3}.
\end{equation}
Putting Eq.\ \ref{eq:pdetect} and \ref{eq:ptransit} together, the
probability that Kepler will detect a planet of radius $R$ orbiting
its host star at period $P$ is 
\begin{equation}
  p_\mathrm{select} = p_\mathrm{transit} p_\mathrm{detect}.
\end{equation}

@@

In addition to true planetary signals, we model a false-positive
background of planet candidates empirically, assuming they populate
the candidate $P$-$R$ plane with a number density that has a linear
gradient across a rectangular region in the $\ln P$-$\ln R$ plane:
\begin{equation}
  \label{eq:background-rate}
  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Nbg}{\Delta \ln P \Delta
    \ln R} \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
    R - \ln R_\mathrm{mid} \right] \right),
\end{equation}
where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$, $\ln
P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
\right)$, $\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
$\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
\right)$.  $\Nbg$, the expected number of background false-positive
events; $P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
$R_\mathrm{min}$, the boundaries in the $P$-$R$ plane within which
background events appear; and $\gamma$, the gradient in the number
density of background events, are parameters of our model.  This is a
purely empirical model for the background contamination, but is
reasonable if the chief contaminant is background eclipsing
binaries\cite{Fressin2013,Duquennoy1991}.

@@

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{selection}
  \caption{\label{fig:selection} \textbf{Comparison of synthetic data
      sets produced from the forward model incorporating selection
      effects with observed candidates.}  (Upper Left) The observed
    (black curve) and synthetic (blue curve; including planets and
    false positives, and using the fitted selection model to
    down-select the candidates from the planet distribution)
    normalised candidate density per logarithmic radius.  Except for a
    discrepancy at $R \simeq 10 \REarth$---associated with hot
    Jupiters, a distinct planetary
    population\cite{Albrecht2012,Naoz2012}---the model produces a good
    fit to the observed candidates over the range of reported radii.
    Note particularly the tail at large radii that comes from
    background contaminants in both observed and synthetic data.
    (Upper Right) Scatter plot of the observed candidates (black
    circles) and a posterior-averaged draw of observed candidates from
    the model (blue circles).  (Lower Left) Scatter plot of the
    observed candidates.  Colors indicate the posterior-averaged
    selection probability for each planet about its host star.  
    The selection probability is treated a product of a geometric
    factor giving the probability of an isotropically-oriented orbit
    producing a transit and a signal-to-noise-ratio-dependent transit
    detection probability.  (Lower Right) The observed (black curve)
    and synthetic (blue curve; including planets and false positives,
    and using the fitted selection model to down-select the candidates
    from the planet distribution) normalised candidate density per
    logarithmic period.  Except for the aforementioned hot Jupiter
    peak at $P \simeq 1 \mathrm{day}$ the model produces a good fit to
    the observed candidates over the range of reported periods.}
\end{figure}

@@

\begin{addendum}
\item The code implementing this analysis is available under an
  open-source ``MIT'' license at
  \url{https://github.com/farr/kepler-selection}.  This work was
  supported by the Science and Technology Facilities Council.
  Computations in this work were performed on the University of
  Birmingham's BlueBEAR cluster.  Some of the data presented in
  this paper were obtained from the Mikulski Archive for Space
  Telescopes (MAST). STScI is operated by the Association of
  Universities for Research in Astronomy, Inc., under NASA contract
  NAS5-26555. Support for MAST for non-HST data is provided by the
  NASA Office of Space Science via grant NNX13AC07G and by other
  grants and contracts.  This paper includes data collected by the
  Kepler mission. Funding for the Kepler mission is provided by the
  NASA Science Mission directorate.
\item [Author Contributions] All authors assisted in the computational
  modelling, discussed the results, and edited the manuscript.
\item [Reprints] Reprints and permissions information is available at
  \url{www.nature.com/reprints}.
\item[Competing Interests] The authors declare that they have no
  competing financial interests.
\item[Correspondence] Correspondence and requests for materials should
  be addressed to W.M.F.\ (email: \email{w.farr@bham.ac.uk}).
\end{addendum}

>>>>

\documentclass[12pt]{article}

\usepackage{scicite}
\usepackage{times}

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

\renewcommand\refname{References and Notes}

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

@@

\bibliographystyle{Science}

@@

%\input{quantities}
\newcommand{\earange}{3.9_{-1.6}^{+2.2}\%}
\newcommand{\rplrange}{3.83_{-0.62}^{+0.76}}
\newcommand{\rpeakrange}{1.25_{-0.17}^{+0.16}}
\newcommand{\corrcoeffrange}{0.334_{-0.053}^{+0.052}}
\newcommand{\fposrange}{7.8_{-1.3}^{+1.4}\%}
\newcommand{\ppeakrange}{0.075_{-0.006}^{+0.007}}
\newcommand{\rhominrange}{5.46^{+0.18}_{-0.18}}
\newcommand{\rhomaxrange}{18.8^{+1.9}_{-1.9}}

\title{The Occurrence of Earth-Like Planets Around Other Stars}
\author{Will M. Farr$^{1\ast}$, Ilya Mandel$^{1}$, Chris Aldridge$^{1}$ \& Kirsty Stroud$^{1}$\\
  \normalsize{$^1$School of Physics and Astronomy}\\
  \normalsize{University of Birmingham}\\
  \normalsize{Birmingham}\\
  \normalsize{B15 2TT}\\
  \normalsize{United Kingdom}\\
  \normalsize{$^\ast$To whom correspondence should be addressed; E-mail: w.farr@bham.ac.uk.}}

\date{}

\begin{document}

\baselineskip24pt

@@

\begin{sciabstract}
  The quantity $\etaearth$, the number density of planets per star per
  logarithmic planetary radius per logarithmic orbital period at one
  Earth radius and one year period, describes the occurrence of
  Earth-like extrasolar planets.  Here we present a
  measurement of $\etaearth$ from a parameterised forward model of the
  (correlated) period-radius distribution and the observational
  selection function in the most recent (Q17) data release from the
  Kepler satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  We find
  $\etaearth = \earange$ (90\% CL).  We conclude that each star hosts
  $\rplrange$ planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim
  0.2 \REarth$.  Our empirical model for false-positive contamination
  is consistent with the dominant source being background eclipsing
  binary stars\cite{Fressin2013}.  The distribution of planets we
  infer is consistent with a highly-stochastic planet formation
  process producing many correlated, fractional changes in planet
  sizes and orbits.  
\end{sciabstract}

The quantity $\etaearth$, the number density of planets per star per
logarithmic planetary radius per logarithmic orbital period at one
Earth radius and one year period, describes the occurrence of
Earth-like extrasolar planets.  Measurement of $\etaearth$ is
complicated by the difficulty of detecting Earth-like planets in
Earth-like orbits about Sun-like stars.  Here we present a measurement
of $\etaearth$ from a parameterised forward model of the (correlated)
period-radius distribution and the observational selection function in
the most recent (Q17) data release from the Kepler
satellite\cite{Borucki2010,Borucki2011,Batalha2013}.  Our data set
comprises 181,568 systems observed under the Kepler exoplanet
observing program (mostly G-type stars on the main
sequence\cite{Batalha2010}), producing 2598 planetary candidates.  We
parameterise the distribution of planetary periods and radii using a
single, correlated Gaussian component; treat selection effects using a
parameterised transit detection probability based on the measured
noise level and stellar properties in the Kepler catalog; and include
an empirically-parameterised, independent component in the
period-radius distribution to represent false-positive planet
detections.  Using our model we can simultaneously estimate
$\etaearth$, place constraints on the planet period-radius
distribution function, and determine the degree of contamination by
false-positive candidate identifications.  We find $\etaearth =
\earange$ (90\% CL).  We conclude that each star hosts $\rplrange$
planets with $P \lesssim 3 \mathrm{yr}$ and $R \gtrsim 0.2 \REarth$,
that the peak of the planet radius distribution lies at $\Rpeak =
\rpeakrange \REarth$, and that $\ln P$ and $\ln R$ are correlated with
correlation coefficient $r = \corrcoeffrange$ (all 90\% CL).  Our
empirical model for false-positive contamination is consistent with
the dominant source being background eclipsing binary
stars\cite{Fressin2013}, with $\fposrange$ (90\% CL) of the candidates
being false-positives.  The distribution of planets we infer is
consistent with a highly-stochastic planet formation process producing
many correlated, fractional changes in planet sizes and orbits.  Our
approach of determining both the intrinsic distribution of objects and
selection effects empirically from survey data is generally
applicable.

@@

The probability that a planet's orbital plane will align with the line-of-sight to
Earth and thereby produce a transit signal is
\begin{equation}
  \label{eq:ptransit}
  p_\mathrm{transit} = 0.0016\, \frac{\RStar}{\RSun}
  \left(\frac{\MStar}{\MSun}\right)^{-1/3} \left(\frac{P}{1\,\mathrm{yr}}\right)^{-2/3}.
\end{equation}
Putting Eq.\ \ref{eq:pdetect} and \ref{eq:ptransit} together, the
probability that Kepler will detect a planet of radius $R$ orbiting
its host star at period $P$ is 
\begin{equation}
  \label{eq:pselect}
  p_\mathrm{select} = p_\mathrm{transit} p_\mathrm{detect}.
\end{equation}

@@

In addition to true planetary signals, we model a false-positive
background of planet candidates empirically, assuming they populate
the candidate $P$-$R$ plane with a number density that has a linear
gradient across a rectangular region in the $\ln P$-$\ln R$ plane:
\begin{equation}
  \label{eq:background-rate}
  \frac{d\Nbg}{d \ln P \, d \ln R} = \frac{\Nbg}{\Delta \ln P \Delta
    \ln R} \left( 1 + \vec{\gamma} \cdot \left[ \ln P - \ln P_\mathrm{mid} , \ln
    R - \ln R_\mathrm{mid} \right] \right),
\end{equation}
where $\Delta \ln P = \ln P_\mathrm{max} - \ln P_\mathrm{min}$, $\ln
P_\mathrm{mid} = 1/2\left(\ln P_\mathrm{max} - \ln P_\mathrm{min}
\right)$, $\Delta \ln R = \ln R_\mathrm{max} - \ln R_\mathrm{min}$,
$\ln R_\mathrm{mid} = 1/2\left(\ln R_\mathrm{max} - \ln R_\mathrm{min}
\right)$.  $\Nbg$, the expected number of background false-positive
events; $P_\mathrm{max}$, $P_\mathrm{min}$, $R_\mathrm{max}$, and
$R_\mathrm{min}$, the boundaries in the $P$-$R$ plane within which
background events appear; and $\gamma$, the gradient in the number
density of background events, are parameters of our model.  This is a
purely empirical model for the background contamination, but is
reasonable if the chief contaminant is background eclipsing
binaries\cite{Fressin2013,Duquennoy1991}.  The posterior on the
background number density in the $P$-$R$ plane appears in
Figure\ \ref{fig:det-bg}.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{selection}
  \caption{\label{fig:selection} \textbf{Comparison of synthetic data
      sets produced from the forward model incorporating selection
      effects with observed candidates.}  (Upper Left) The observed
    (black curve) and synthetic (blue curve) normalised candidate
    density per logarithmic radius.  Except for a discrepancy at $R
    \simeq 10 \REarth$---associated with hot Jupiters, a distinct
    planetary population\cite{Albrecht2012,Naoz2012}---the model
    produces a good fit to the observed candidates over the range of
    reported radii.  Note particularly the tail at large radii that
    comes from background contaminants in both observed and synthetic
    data.  (Upper Right) Scatter plot of the observed candidates
    (black circles) and a posterior-averaged draw of observed
    candidates from the model (blue circles).  (Lower Left) Scatter
    plot of the observed candidates.  Colors indicate the
    posterior-averaged selection probability for each planet about its
    host star (see Eq.\ \eqref{eq:pselect}).  (Lower Right) The
    observed (black curve) and synthetic (blue curve) normalised
    candidate density per logarithmic period.  Except for the
    aforementioned hot Jupiter peak at $P \simeq 1 \mathrm{day}$ the
    model produces a good fit to the observed candidates over the
    range of reported periods.}
\end{figure}

@@

@@

\begin{scilastnote}
\item The code implementing this analysis is available under an
  open-source ``MIT'' license at
  \url{https://github.com/farr/kepler-selection}.  This work was
  supported by the Science and Technology Facilities Council.
  Computations in this work were performed on the University of
  Birmingham's BlueBEAR cluster.  Some of the data presented in
  this paper were obtained from the Mikulski Archive for Space
  Telescopes (MAST). STScI is operated by the Association of
  Universities for Research in Astronomy, Inc., under NASA contract
  NAS5-26555. Support for MAST for non-HST data is provided by the
  NASA Office of Space Science via grant NNX13AC07G and by other
  grants and contracts.  This paper includes data collected by the
  Kepler mission. Funding for the Kepler mission is provided by the
  NASA Science Mission directorate.
\end{scilastnote}
